---
title: "Relational data"
author: "attackgnome"
date: "10/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 13.2.1 Exercises

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

  We would need the latitude and longitude of the origin and destination airport. We would need to combine the information in the airports and flights table.

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

  It would be a line connecting the `faa` field in the airports table to the `origin` field in the weather table.

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

  There would also be a relation to the destination field in the flights table

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

  One way to represent it would be a table consisting of the month and day of the "special" day and a comment describing what the day is. The primary keys would be the month and day and would allow the table to be linked to the flights table.
  
  
## 13.3.1 Exercises
1. Add a surrogate key to flights.

```{r exercises_3.1.1}
flights %>%
  mutate(key = row_number())
```

2. Identify the keys in the following datasets
  a. Lahman::Batting
  `playerID`, `yearID`, `teamID`, `lgID`
  
  b. babynames::babynames
  None - no other tables
  
  c. nasaweather::atmos
  `lat`, `long`
  
  d. fueleconomy::vehicles
  `make`, `model`
  
  e. ggplot2::diamonds
  None - no other tables
  
  (You might need to install some packages and read some documentation.)

3. Draw a diagram illustrating the connections between the Batting, People, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between People, Managers, AwardsManagers.

4. How would you characterise the relationship between the Batting, Pitching, and Fielding tables?  


## 13.4.6 Exercises
1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r exercises_4.6.1}
delay <- flights %>%
  group_by(dest)%>%
  summarize(delay = abs(mean(arr_delay, na.rm = TRUE))+ abs(mean(dep_delay, na.rm = TRUE)))

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  left_join(delay, c("faa" = "dest"))%>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(colour = delay)) +
    coord_quickmap()
```


2. Add the location of the origin and destination (i.e. the lat and lon) to flights.


```{r exercise_4.6.2}
airport_loc <- airports%>%
  select(lat, lon, faa)

flights %>%
  left_join(airport_loc, c("origin" = "faa")) %>%
  rename(origin_lat = lat, origin_lon = lon) %>%
  left_join(airport_loc, c("dest" = "faa")) %>%
  rename(dest_lat = lat, dest_lon = lon)
```


3. Is there a relationship between the age of a plane and its delays?

```{r exercise_4.6.3}
planes %>%
  select(tailnum, model, plane_year = year)%>%
  inner_join(flights, "tailnum") %>%
  group_by(plane_year)%>%
  summarise(avg_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot() + geom_point(aes(x=plane_year, y = avg_delay))
```


4. What weather conditions make it more likely to see a delay?

```{r exercise_4.6.4}
flights_weather <- flights %>%
  inner_join(weather, by = c("origin"= "origin", "year" = "year", "month" = "month", "day" = "day", "hour"= "hour"))

#Precipitation
flights_weather%>%
  group_by(precip)%>%
  summarise(delay = mean(dep_delay, na.rm = TRUE))%>%
  ggplot(aes(x=precip, y=delay))+ geom_line() + geom_point()

#Visibility
flights_weather%>%
  mutate(visib_cat = cut_interval(visib, n = 10)) %>%
  group_by(visib_cat) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = visib_cat, y = dep_delay)) +  geom_point()

#Temp
flights_weather%>%
  mutate(temp_cat = cut_interval(temp, n = 10)) %>%
  group_by(temp_cat) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = temp_cat, y = dep_delay)) +  geom_point()

```


5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

```{r exercise_4.6.5}
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() 
```


## 13.5.1 Exercises
1. What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

It means the flight never departed, (i.e. the flight was cancelled)

```{r exercise_5.1.1}
flights %>% filter(is.na(tailnum))
```


2. Filter flights to only show flights with planes that have flown at least 100 flights.

```{r exercise_5.1.2}
flights %>% 
  right_join(planes, "tailnum") %>%
  group_by(model) %>%
  summarise(n = n())%>%
  filter(n>99) %>%
  arrange(desc(n))
```

3. Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

```{r exercise_5.1.3}
vehicles %>% 
  right_join(common, c("make", "model"))%>%
  group_by(make)%>%
  count()
```


4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

```{r exercise_5.1.4}
worst_hours <- flights %>%
  mutate(hour = sched_dep_time %/% 100) %>%
  group_by(origin, year, month, day, hour) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(dep_delay)) %>%
  slice(1:48)

weather_most_delayed <- semi_join(weather, worst_hours, 
                                  by = c("origin", "year",
                                         "month", "day", "hour"))

select(weather_most_delayed, temp, wind_speed, precip) %>%
  print(n = 48)

ggplot(weather_most_delayed, aes(x = precip, y = wind_speed, color = temp)) +
  geom_point()
```

5. What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

```{r exercise_5.1.4}
#Drops all observations in flights where "dest" matches airports$faa
anti_join(flights, airports, by = c("dest" = "faa"))

#Drops all observations in airports where "faa" matches flights$dest
anti_join(airports, flights, by = c("faa" = "dest")) 
```

6. You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

```{r exercise_5.1.4}
flights %>%
  filter(!is.na(tailnum)) %>%
  distinct(tailnum, carrier, .keep_all = TRUE)%>%
  count(tailnum) %>%
  filter(n > 1) %>%
  inner_join(flights, "tailnum") %>%
  select(carrier, tailnum, n) %>%
  # join with airlines to get airline names
  left_join(airlines, by = "carrier") %>%
  select(tailnum, carrier) %>%
    inner_join(airlines, "carrier") %>% unique()

```
