---
title: "Relational data"
author: "attackgnome"
date: "10/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 13.2.1 Exercises

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

  We would need the latitude and longitude of the origin and destination airport. We would need to combine the information in the airports and flights table.

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

  It would be a line connecting the `faa` field in the airports table to the `origin` field in the weather table.

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

  There would also be a relation to the destination field in the flights table

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

  One way to represent it would be a table consisting of the month and day of the "special" day and a comment describing what the day is. The primary keys would be the month and day and would allow the table to be linked to the flights table.
  
  
## 13.3.1 Exercises
1. Add a surrogate key to flights.

2. Identify the keys in the following datasets
  a. Lahman::Batting,
  b. babynames::babynames
  c. nasaweather::atmos
  d. fueleconomy::vehicles
  e. ggplot2::diamonds
  (You might need to install some packages and read some documentation.)

3. Draw a diagram illustrating the connections between the Batting, People, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between People, Managers, AwardsManagers.

4. How would you characterise the relationship between the Batting, Pitching, and Fielding tables?  

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
